What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts

### 2023-10-27 revisions-round1/response-to-reviewers.qmd

    -   *In response to specific comments from Reviewer 2, we have strengthened some of the wording in the introduction to better emphasize the importance of our work and the survey framework. We hope this addresses this concern.*

-   The entire article reads more like a project report rather than a scientific paper, with an excess of questions and bullet-pointed sentences.

    -   *To address this, we have removed two of the instances of bulleted lists: one which listed out the three variations in testing, each described in a bullet, and a second which demonstrated the options participants have to select from, each option shown in a bullet. The third list, which we have kept intact, is a list of the three research questions we pose to answer through our study. The authors could not identify a reason to remove the list structure for our research questions, as it clearly separates each of the distinct questions for the reader. There are four questions in the text of the paper: one at the very beginning, posing the question 'What do viewers see when we show them a data chart?', and three associated with the research questions. We have not removed any of these, as a question mark at the end of a research question is appropriate, and posing an opening question in the introduction is also appropriate.*

-   There is an overabundance of figures without clear emphasis.

    -   *We have removed a redundant sub-figure from Figure 4. Displaying the response patterns for each research question visually is a useful method of communicating our findings in addition to formal modeling results. The inclusion of our results figures highlights the nuance in our findings, particularly when it comes to comparisons across tasks and subgroups, and we feel they are important to convey our findings. For that reason, we have retained most of them.*

@@

        -   page 2, third paragraph: "to be biased towards more male, younger and relatively higher education" -\> "to be biased towards more male, younger and relatively well-educated individuals"
            -   *This change has been made in the text.*

        -   page 2, third paragraph: "Further, the populations' emphasis..." -\> This sentence was confusing. I kind of get the gist what the authors are trying to say but I wonder if it can be paraphrased better? E.g. "Arguably, we could limit the population of interest to highly educated individuals, however, this group is also likely to have prior exposure to data analysis in higher education, which may not be the case for other groups in the wider population."
            -   *We have rephrased this section to make the point more clearly.*

        -   page 2, fourth paragraph: I"m not sure if you should start the paragraph with "Here".
            -   *We have replaced 'Here' with 'In this work'*

    -   Figure 1, the text within the visualisation are too small to see. Perhaps lay it out as a 3 x 2 instead?
            - *Figure 1 was not meant to have readable text within the stimuli visualizations but only give an overview of the different stimuli and the testing situations. We have increased the font size for the overview, expanded the discussion of the figure to reflect its purpose and added image (a) to the supplement in a larger format.*

@@

    -   page 8, first paragraph: "\$ 2\$" is a typo?
        -   *This was a typesetting error and has been resolved in the source code.*

    -   Table 1, effective sample size adds up to 1520.3 but the total says 1520.8? 
        -   *What is shown is the effective sample size of the combined survey rounds, not the total of each round's effective sample size. The fact that these numbers are close is due to similar weighting strategies being used in each of the individual rounds. We have now renamed the line "Combined Rounds" to reduce confusion.*

    -   Figure 3: I take it that the American Community Survey 2021 is considered the Census estimates? I think it makes more sense to swap the bar and point so that the bar is the Census estimate. This way you can use the errorbar plot. The margin of error around a barplot is reminiscent of a dynamite plot, which I don't think is well thought of in the community (at least not me). I think you can also color the bars to indicate whether your survey estimate covers the census estimate so it is easier to see which one covers it or not. In this plot, you've facetted by the stimulus type but I think it makes sense to facet by the category level (the comparison of distribution is done with the census estimate directly, and I think the interest is to compare the distribution across stimulus so you want the primary factor of interest to compare to be close in proximity). Within a category level, the census estimate is constant across stimulus so you could substitute is with a horizontal line (instead of a bar or point) as well.
        -   *We have changed the format of this figure to address these concerns, using a dot plot with error bars to represent the uncertainty.*

    -   Figure 4: I think (b) alone is sufficient as you can gain the same information as (a) (although perhaps it is less familiar for wider audience). You could arguably put figure (a) information in a table instead.
        -   *We have removed Figure 4(a) and just kept what was originally 4(b) to avoid this redundancy.*

@@

## Editors

-   The AE and I do believe this is within the scope for this special issue, contrary to the final comment from Referee #1, so no need to address that comment.

-   Please add an unnumbered section Supplementary Material to list its content (e.g., data, code, additional simulation results, technical proofs, etc.). See recently published papers for examples. Reproducibility is held high at the JDS; the reproducibility team will check for reproducibility when the paper is accepted.

    -   *We have added this section to the end of the paper, and note that we also submitted a supplement document.*

@@

-   When addressing the comments, please keep in mind that there is a 20-page limit (including everything) in the production style.

-   In general, both reviewers express some concerns regarding some of the language and description regarding overall impact of the work. I would encourage the authors to make it more clear that their primary contribution is in using a representative probability sample, and that their conclusions are focused on stacked barplot displays, as Reviewer 2 specifically remarks, and highlight why conclusions from a non-representative sample may be problematic within the data visualization world.

    -   *We have updated the language in the title, abstract, introduction, and conclusion to be responsive to these concerns. See our specific responses to Reviewer 2's specific comments, but in general we added more emphasis on the importance of communicating a general audience, why non-representative survey findings don't necessarily hold for the entire population, and clarified that our scope was specifically on stacked bar chart displays.*

-   I believe the figures are fairly clear (per Reviewer 1's comment), but the authors might also be inclined to be more selective with what is shown in Figures 3, 6-8, and maybe move some of the others to an appendix.

    -   *In response to a specific comment from Reviewer 2, we have removed a redundant sub-figure from Figure 4. We have retained the other figures given that we are still within the page limit and displaying the response patterns visually is a useful method of communicating our findings in addition to formal modeling results. The inclusion of our results figures highlights the nuance in our findings, particularly when it comes to comparisons across tasks and subgroups.*

### 2023-10-27 SDSS-2023/testing-charts-accuracy.qmd
figure 4

Cleveland and McGill's work -- while a foundational user study in graphical perception -- utilized a small convenience sample, consisting of only a few individuals recruited from among the authors' coworkers and their spouses.
@heerCrowdsourcingGraphicalPerception2010 reproduced Cleveland and McGill's rankings using a larger sample from a crowd sourcing platform, employing a total of 82 Amazon Mechanical Turk (mTurk) workers for the study.
Crowd sourced samples were shown by @borgoCrowdsourcingInformationVisualization2017 to be biased towards more male, younger, and relatively well-educated individuals relative to the adult U.S. population as a whole; generally, convenience and opt-in samples are not representative of the general population.
Further, higher education individuals are likely to have prior exposure to data visualization or data topics in higher education, which may not be the case for other groups within the population; existing differences in exposure to data may drive differences in results across groups. The general public is a common target audience for data visualization and scientific communication work; results of polls, surveys, and scientific work are presented often in the news, shared on social media, or posted online or in print materials. It is imperative that data visualizations used to communicate these topics are designed with the goal of reaching the entire population. To that end, an inclusive, representative approach to studying how viewers perceive information presented in charts is critical to understand how best to communicate with the general public.

In this work, we seek to answer whether it is possible to reproduce some of the previous findings in the context of a survey with a large, nationally-representative set of respondents.
Specifically, we present viewers with structural variations of bar charts and ask them to answer questions comparing the size of elements within those charts.
We employ a probability-based survey panel and run a series of perception tests with nationally-representative samples of respondents from that panel.
The advantage of using a probability-based approach is two-fold.
First, we have access to a large sample of survey participants and thus have greater power in making inference about graphical perceptional abilities.
Second, the sample is representative of the general adult public in the U.S., and this allows us to test whether prior results from convenience samples hold with a nationally representative sample and whether there are differences in those results across demographic subgroups.
Within that context we focus on the following research questions:

@@

age <- acs_age %>%
  ggplot(aes(x = age4)) +
  geom_errorbar(aes(ymin = percent2-moe_percent, ymax=percent2+moe_percent, colour="ACS estimate")) +
#  geom_point(aes(y = percent)) +
  geom_point(
    size=3,
    aes(x = age, y = value, colour=chart),
    position = position_dodge(width = 0.5),
    data = omni_age %>% filter(group=="mean")
  ) +
  ylim(c(0, NA)) +
  geom_linerange(
    aes(
      x = age,
      ymin = low,
      ymax=high,
      colour=chart
    ),
    linewidth = 1,
    position = position_dodge(width = 0.5),
    data = omni_age %>% pivot_wider(names_from=group, values_from=value)
  ) +
  ylab("Percent") + 
  ggtitle("Age groups") +
  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) +
  xlab("") +
  theme(legend.position ="none") +
  scale_linetype_manual(values=c(2,2,1)) +
  scale_color_manual(values=c("grey20", rounds_col))

@@

gender <- 
  gender +
  annotate("text", x = 0.83, y = 5, label="ACS estimate", hjust=0, 
           size = 5, colour = "grey20") +
  annotate("segment", x = 0.5, xend=0.7, y = 5,yend=5, 
           linewidth = 1, colour="grey20") +
  annotate("text", x = 0.83, y = 15, label="Horizontal design", hjust=0,
           size = 5, colour = "grey20") +
  annotate("segment", x = 0.8333, xend = 0.8333, y = 17, yend=40, 
           linewidth = 0.75, colour=rounds_col[1], linetype=2) +
  annotate("text", x = 1, y = 20, label="Horizontal wide", hjust=0,
           size = 5, colour = "grey20") +
  annotate("segment", x = 1, xend = 1, y = 22, yend=40, 
           linewidth = 0.75, colour=rounds_col[2], linetype=2) +
  annotate("text", x = 1.17, y = 25, label="Vertical design", hjust=0,
           size = 5, colour = "grey20") +
  annotate("segment", x = 1.166, xend = 1.166, y = 27, yend=40, 
           linewidth = 0.75, colour=rounds_col[3], linetype=2)

@@

\autoref{fig-alignment}(a) displays all responses along the binary correctness measure, separated by whether the stimulus was an aligned task or unaligned task.
We can see that levels of accuracy for all responses are significantly higher for the 'easier' (aligned) task, with about twice as many respondents correctly selecting the larger of the two marked elements.

@@

```{r accuracy,  warning=FALSE}
#| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (green) and incorrect (grey) responses to the two comparison questions. When tiles are aligned along the same axis, more than twice the number of responses are correct. The shaded area along the top of the green bars corresponds to  95\\% confidence intervals around (marginal) correct responses. On the right (b), a parallel coordinate plot shows all combinations of responses. There's a huge asymmetry in the number of responses where participants answered only one of the questions correctly."
#| label: fig-alignment
#| out-width: 90%
#| fig-width: 8.5

@@

gg1 + gg2
```

@@

>>>>

Cleveland and McGill's work -- while a foundational user study in graphical perception -- utilized a small convenience sample, consisting of only a few individuals recruited from among the authors' coworkers and their spouses.
@heerCrowdsourcingGraphicalPerception2010 reproduced Cleveland and McGill's rankings using a larger sample from a crowd sourcing platform, employing a total of 82 Amazon Mechanical Turk (mTurk) workers for the study.
Crowd sourced samples were shown by @borgoCrowdsourcingInformationVisualization2017 to be biased towards more male, younger, and relatively well-educated individuals compared to the general adult U.S. population. 
Presumably, individuals with a higher education are more likely to have prior exposure to data visualization and might be more familiar with quantitative statements than less well-educated groups within the population; other pre-existing differences between population groups may also drive differences in understanding or interpreting charts. 
This raises the question of whether conclusions drawn from these studies hold for the general population -- 
<!--generally, convenience and opt-in samples are not representative of the general population.-->
 a common target audience for data visualization and scientific communication work. Results of polls, surveys, and scientific work are presented visually in the news, shared on social media, posted online or in print materials. It is imperative that data visualizations used to communicate quantitative information are designed with the goal of reaching the entire population. To that end, an inclusive, representative approach to studying how viewers perceive information presented in charts is critical to understand how best to communicate with the general public.

In this work, we seek to answer whether it is possible to reproduce some of the previous findings in the context of a survey with a large, nationally-representative set of respondents.
Specifically, we present viewers with structural variations of bar charts and ask them to answer questions comparing the size of elements within those charts.
We employ a probability-based survey panel and run a series of perception tests with nationally-representative samples of respondents from that panel.
The advantage of using a probability-based approach is two-fold:
First, the relatively larger sample of survey participants provides greater statistical power in drawing inference about graphical perceptional abilities.
Second, the sample is representative of the general adult public in the U.S., permitting an assessment inhowfar prior results from convenience samples hold with a nationally representative sample and whether there are differences in those results across demographic subgroups.
Within that context we focus on the following research questions:

@@

age <- acs_age %>%
  ggplot(aes(x = age4)) +
  geom_errorbar(aes(ymin = percent2-moe_percent, ymax=percent2+moe_percent, colour=" ACS estimate")) +
#  geom_point(aes(y = percent)) +
  geom_point(
    size=3,
    aes(x = age, y = value, colour=chart),
    position = position_dodge(width = 0.5),
    data = omni_age %>% filter(group=="mean")
  ) +
  ylim(c(0, NA)) +
  geom_linerange(
    aes(
      x = age,
      ymin = low,
      ymax=high,
      colour=chart
    ),
    linewidth = 1,
    position = position_dodge(width = 0.5),
    data = omni_age %>% pivot_wider(names_from=group, values_from=value)
  ) +
  ylab("Percent") + 
  ggtitle("Age groups") +
  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) +
  xlab("") +
  theme(legend.position ="none") +
  scale_linetype_manual(values=c(2,2,1)) +
  scale_color_manual(values=c("grey20", rounds_col))

@@

gender <- 
  gender +
  annotate("text", x = 0.83, y = 5, label="ACS estimate & 95% C.I.", hjust=0, 
           size = 5, colour = "grey20") +
  annotate("segment", x = 0.5, xend=0.7, y = 5,yend=5, 
           linewidth = 1, colour="grey20") +
  annotate("segment", x = 0.5, xend=0.7, y = 5.1,yend=5.1, 
           linewidth = 1, colour="grey20") +
  annotate("segment", x = 0.5, xend=0.7, y = 4.9,yend=4.9, 
           linewidth = 1, colour="grey20") +
  annotate("text", x = 0.83, y = 15, label="Horizontal design", hjust=0,
           size = 5, colour = "grey20") +
  annotate("segment", x = 0.8333, xend = 0.8333, y = 17, yend=40, 
           linewidth = 0.75, colour=rounds_col[1], linetype=2) +
  annotate("text", x = 1, y = 20, label="Horizontal wide", hjust=0,
           size = 5, colour = "grey20") +
  annotate("segment", x = 1, xend = 1, y = 22, yend=40, 
           linewidth = 0.75, colour=rounds_col[2], linetype=2) +
  annotate("text", x = 1.17, y = 25, label="Vertical design", hjust=0,
           size = 5, colour = "grey20") +
  annotate("segment", x = 1.166, xend = 1.166, y = 27, yend=40, 
           linewidth = 0.75, colour=rounds_col[3], linetype=2)

@@

\autoref{fig-alignment} displays all responses along the binary correctness measure, separated by whether the stimulus was an aligned task or unaligned task.
We can see that levels of accuracy for all responses are significantly higher for the 'easier' (aligned) task, with about twice as many respondents correctly selecting the larger of the two marked elements.

@@

```{r accuracy,  warning=FALSE}
#| fig-cap: "Participants' response pattern to both questions. The coloring reflects the number of correct responses (two in full green, zero in grey, and one correct answer in grey-green). When tiles are aligned along the same axis, more than twice the number of respondents correctly identifies the larger tile. There's a huge asymmetry in the number of responses where participants answered only one of the questions correctly."
#| label: fig-alignment
#| out-width: 40%
#| fig-width: 4

@@

gg2
```

@@

# Supplementary Material

The source code and data used in this paper are available on a GitHub repository at [https://github.com/kiegan/testing-charts-jds](https://github.com/kiegan/testing-charts-jds). In addition, we provide an example stimulus image in that repository.
